---
title: "Statistical Inferance Course Project Question 1"
author: "Sheetal"
date: "13 October 2018"
output: pdf_document
---

#Simulation Exercise
The purpose of this project is to investigate the exponential distribution in R and compare it with the Cental Limit Theorem.
The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter.
 compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. 
 
 Lambda = 0.2
 Simulations = 1000



#Question 1
Show the sample mean and compare it to the theoretical mean of the distribution.

```{r}
lambda <- 0.2
Data <- matrix(rexp(1000*40, lambda), nrow = 1000, ncol = 40)
distMean <- apply(Data, 1, mean)
```

```{r echo=TRUE, fig.width=7, fig.height=3, fig.align='center'}

hist(distMean, breaks = 50, main = "The distribution of 1000 averages of 40 random exponentials", 
      xlab = "Value of means", ylab = "Frequency of means", col = "pink")
abline(v = 1/lambda, lty = 1, lwd = 5, col = "yellow")
legend("topright", lty = 1, lwd = 5, col = "yellow", legend = "theoretical mean")

```

In the abopve chart, I have simulated the exponential distribution 1000 times, and created a distrbution of the means. The AB line is at the mean. The results show that the sample mean is close to the theoretrical mean.


the below is the mean for all the sumulations one at a time #
```{r}
simMean <- rowMeans(Data)
```

the below is the mean for the data, in this case 4.9632
```{r}
simMeanMean <- mean(simMean)
```

the theatrical mean is 1/lambda i.e 1/0.2, in this case 5
```{r}
theatricalmean <- 1/0.2
```
both the sample mean and the theatrical mean are similar


#Question 2
Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
```{r}
simulationvar <- var(simMean)
theatricalvar <- (1/lambda)^2/(40) 
```

simulation variance is 0.61, theatrical variance is 0.62 and hence they are close

#Question 3
Show that the distribution is approximately normal
```{r echo=TRUE, fig.width=7, fig.height=3, fig.align='center'}
hist(distMean, breaks = 50, main = "The distribution of 1000 averages of 40 random exponentials", 
     xlab = "Value of means", ylab = "Frequency of means", col = "pink")
```

The output graph looks close to a normal distribution