---
title: "Regression Models Course Project"
author: "Sheetal"
date: "19 October 2018"
output: pdf_document
---


##Executive Summary'
The purpose of this report is the answer 2 questions :
1.“Is an automatic or manual transmission better for MPG”
2."Quantify the MPG difference between automatic and manual transmission

Information on the datasets is found in the R help guides. There are 10 variables :
1 mpg: Miles/(US) gallon
2 cyl: Number of cylinders
3 disp: Displacement (cu.in.)
4 hp: Gross horsepower
5 drat: Rear axle ratio
6 wt: Weight (1000 lbs)
7 qsec: 1/4 mile time
8 vs: V/S
9 am: Transmission (0 = automatic, 1 = manual)
10 gear: Number of forward gears
11 carb: Number of carburetors


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Expoloratory Data Analysis
The below sections peforms some preliminary exporation of the dataset

```{r,echo=TRUE}
mtcars <-mtcars
summary(mtcars)
```
There ae 32 observations.

The summary statistics shows that there is 1 dependendant variable and 10 explanotary variables. VS and AM are factor variablese so they need to be catered for 

```{r}
mtcars$vs <- as.factor(mtcars$vs)
mtcars$am <- as.factor(mtcars$am)
```

#1.“Is an automatic or manual transmission better for MPG”


```{r,echo=TRUE}
boxplot(mpg ~ am, data = mtcars,
         col  = c("pink", "blue"),
         xlab = "Transmission",
         ylab = "Miles per Gallon",
         main = "MPG by Transmission Type",
         names= c("automatic","manual"),
         horizontal= T) 

```

From the box plot on appears that the manual tranmission has greater miles per gallon on average.

This can be double checked via hyphothesis testing
            
              H0 : The transmission type has no impact on miles per gallon
              H1 : The transmission type impacts mile per gallon

```{r}
auto=subset(mtcars,select=mpg,am==0)
manual=subset(mtcars,select=mpg,am==1)
t.test(auto,manual)
```

As the p-value is small we can reject the null hypothesis.

#2."Quantify the MPG difference between automatic and manual transmission

#Linear Regressions

```{R}
simpleregression <- lm(mpg~am,mtcars) 
summary(simpleregression)
```

Doing a simple linear regression between miles per gallon (MPG) and transmission, shows that manual is better by 7.245 miles per gallon. The R squared is only 36%, so there is more variance to explain,

The next step is multiple regession

```{R}
multipleregression <- lm(mpg~.,mtcars)
summary(multipleregression)
```

The R squared has now improved to 86.9%,  It now looks like manual is better by 2.52 miles per gallon.The p values have all gone up, and now none of the coeffincents are significant.

I am now going to try step wise regression

```{R}
regSR=step(multipleregression,trace=0)
summary(regSR)

```

The R squared is 85%, and now we are left with significant coefficents :
wt: Weight (1000 lbs)
qsec: 1/4 mile time
Transmission (0 = automatic, 1 = manual)

We can now conclude that manual transmission allows 2.9358 miles than automatic transmission.

```{R}
plot(regSR, which=c(1:1))
```
The residuals look fairly low for the final stepwise model